Objectification in Action: Self- and Other-Objectification in Same-gender Interactions

### 2020-07-21 Final_Manuscript.Rmd
paragraph on authenticity

@robertsfredrickson's Objectification Theory, suggests that women are objectified in interpersonal encounters, and that the effect might be strongest from men, but possibly also from women. Psychological researchers studying the sexual objectification of women, and the resulting self-objectification that follows, have in recent years explored the interpersonal process of the objectification - self-objectification link. For exmaple, in a face-to-face mixed-gender dyadic study, @garcia2016objectification found that men's reported objectificaiton of their female interaction partner during an interaction was associated with increased sefl-objectification reported by their partner. @gervais2020social has reviewed the literature on interpersonal self-objectification and has organized the the process in a theoretical model called the SIMO, but this model is only relevant for mixed-gender encouters. Although there is ample evidence that women objectify other women [@loughnan2015exploring; @puvia2013being], there is currently no studies investigating the process of interpersonal objectification same-gender interactions. Does objectification by a female interaction partner have the same downstream negative consequences as being objectified by a male interaction partner? For example, perhaps women could be amplifying each others' self-objectification as they get ready to go out for the night, *before* the theoretical encounters with men outlined in @gervais2020social occur.

@@

SIMO model and literature review [@gervais2020social]. Studies have shown that within social encounters women are gazed at more than men [@briton1995, CITE saguy skype study, calogero video walking study], often times feel ‘looked at’ within interpersonal interactions [@argyle1969], and will more than likely internalize the objectifying gaze on physical self [@young1979]. Moreover, perhaps the most adverse effect of objectifying treatment is that it effectively socializes girls and women to treat themselves as objects to be looked at and evaluated, an effect termed self-objectification  [@Bartky; @berger1972; @fredricksonetal1998]. We don't know about women-on-women interpersonal objectification. What might be different?

Self-objectification has been found to occur after a mere relationship prime among women [@sanchez2008romance] because in our culture women need to look attractive to obtain and maintain successfuol relationship, thus, the "college relationships condition may highten self-objectification and the evaluation of other women in a sexualized way. Researchers have also studied interpersonal objectification and self-objectification in romantic relationships [@strelan2018birds]. Evidence is mixed, and women can sometimes have positive outcomes from appearance valuation within the context of an established romantic relationship and they experience less SSO after appearance comments within romantic relationships [@meltzer2020women], less is known about first dates. But this literature is also only about heterosexual relationships. A recent study has investigated the effect of self-objectification on the reduction of relationship biulding skills in general (including same-sex friendships) [CITE PWQ paper I reviewed, Yoder editor]. Empirical evidence reveals that objectification manifests through inauthenticity in romantic relationships [@brunelletal2010], adverse attitudes in regard to career aspirations, and a decrease in concentration and impairment in female cognitive performance [@quinnetal].

MAYBE SEPARATE PARAGRAPH ABOUT AUTHENTICITY LINK SPECIFICALLY. And how woman-woman interpersonal objectificaiton preocesses might diverge from this process. [again CITE PWQ paper I reviewed, Yoder editor].

@@

The decision was made to ask even heterosexual women to judge their women partners as potential romantic partners. We felt that this prompt would keep the study closest to a replication of the previous @garcia2016objectification version of the study. Past research has found that women, largely due to an increase self-objectifying state, are indeed able to evaluate other women's potential as romantic partners--indeed, women may be unfortunutely used to thinking this way about themselves, and we suspect other women.

@@

#Smith--external and internal
smith_funcTSO_alpha <- psych::alpha(select(smith_final, tso_1, tso_2, tso_4, tso_7, tso_9), check.keys=TRUE)$total[2]
smith_physTSO_alpha <- psych::alpha(select(smith_final, tso_3, tso_5, tso_6, tso_8, tso_10), check.keys=TRUE)$total[2]
```
Trait self-objectification (TSO) was assessed using the Self-Objectification Questionnaire [@fredricksonetal1998; @nollfredrickson1998], which evaluates the extent to which individuals view their bodies in observable versus nonobservable ways. The questionnaire asked participants to rank order both appearance and functional aspects of their bodies, from 1 (least important) to 10 (most important), with respect to physical self-concepts. Of the ten body attributes, five of the items were appearance-based (weight, sex appeal, physical attractiveness, firm/sculpted muscles and body measurements), and five of the items were competence-based (strength, physical coordination, energy level, health and physical fitness). Difference scores were computed by subtracting the sum of the 5 functional aspects/competence attributes (e.g., health, strength) from the sum of the 5 physical self-concepts/appearance attributes (e.g., physical attractiveness, weight), and all measures were multiplied by -1, as was done in @garcia2016objectification, so that positive scores indicated greater TSO.

@@

```{r 2016figure, include=TRUE, echo=FALSE, results='asis', out.width="400px", fig.cap="Path Analysis Model from Garcia et al. (2016) study with distinguishable dyads."}
include_graphics("2016_figure.jpg")
```
This study sought to replicate the results of @garcia2016objectification’s study, done with male-female pairs, which used a dyadic path analysis to detect whether partners’ objectification of one another affected state self-objectification (SSO). See Figure\ \@ref(fig:2016figure) for the results of the analysis from this previous study. We hope to investigate how the central effects found in the previous study relate to interactions between two women. Specifically, we are interested in testing the relationship between state-other objectification and SSO, and how SSO in turn, affecs feelings of inauthenticity during the interaction. In addition, we will also test if the effect of other-objectification in an interaction on SSO is only present for those women who are high in trait self-objectification, as in @garcia2016objectification. Further, we will investigate the realtionships between experiencing interaction inauthenticity and relationship agency, career aspirations, and cognitive performance.

While @garcia2016objectification used dyadic path analysis, we will conduct our dyadic analyses using multilevel modeling. Dyadic analyses for distingishable dyads (e.g., mixed-gender interacting pairs) is more natural in Structual Equation Modeling (SEM) than it is for indistinguishable dayds (e.g., same-gender interacting pairs) [@ledermann2017analyzing; @garcia2015moderation]. One reason for this asymmetry is that, due to the arbitrary distinctions made bewteen "partner 1" and "partner 2" in indistinguishable dyads, many estimates need to be fixed to be equal (i.e., paths, variances, covariances, endogenous intercepts, and exogenous means) for indistinguishable dyads but these equality contraints should not then be considered in the degrees of freedom calculations for fit estimations [@olsen2006structural]. Further, @olsen2006structural detail how a new independence model and the corresponding fit measure should be re-calculated for indistinguishable dyads models. The current study uses dyadic multilevel modeling (MLM) to test all relationships and mediation paterns. The online supplementary materials contains analysis using SEM. See @ledermann2017analyzing for a more complete discussion of the considerations for using SEM versus MLM for dyadic analysis.

Testing the @garcia2016objectification model on the current, same-gender, sample, involves using the Actor-Partner Independence Model (APIM) approch for each outcome variable (i.e., endogenous variable in Figure\ \@ref(fig:2016figure)). Thus, we ran five APIM's to test all the hypothesized relationships. See Figure\ \@ref(fig:apim) for a basic APIM model. The APIM includes effects due to one's own, as well as one's partner's, predictor variables ($X$'s) on the one's own outcome variable ($Y$). Unlike the original @garcia2016objectification study, our study deals with indistinguishable dyads, meaning the designation of who is designated as "actor" and who is designated as "partner" is arbitrary. Recall that the indistinguiahble nature of the dyads in the current study led us to choose the MLM approach over SEM. These analyses are considered exploratory, given the lack of prior research theorizing about these linkages.

@@

round(corr.test(select(filter(clarksmith_women, smith_or_clark == "Clark"), 
                 a2pSOO_A, a2pSOO_P, TSO_A, 
                 howbody_avg_A,
                 authenticity_A, 
                 trigram_score_A, careerasp_A, relag_A))$r, 3)
```
The correaltions between study variables is similar across samples. The realibialities for the study scales were also equivalent. (Note that the two samples were two small to conduct formal measurement equivalence tests for scales.) There are no statistically significant differences between samples in demographics including age, STATS, and ethnicity, STATS.

@@

summary(apim1)
summary(apim1.b)
```
The most important finding from @garcia2016objectification was the significant partner effect of other objectification and SSO (specifically men's objectification of women and women's SSO). As expected, the partner effect of other-objectification on SSO in the current all-women sample was statistically signficant, `r report_stats(apim1, 3)`, replicating @garcia2016objectification's fiding. One's own other objectification had no effect on SSO, `r report_stats(apim1, 2)`. Contrary to past finding however, there was no statistically significant interaction of partner's other objectification and the person's trait self-objectification on SSO, `r report_stats(apim1, 5)`. There was also no significant main effect of trait self-objectification on SSO `r report_stats(apim1.b, 4)`.

@@

summary(apim2.c)
```
Contrary to expectations, there was no significant effect of SSO on interaction authenticity, although the estimate of this effect was in the hypothesized negative direction, `r report_stats(apim2, 2)`. Because authenticity was a composite score of 9 items, two of which were interaction specific authenticity items, we also estimated the pairwise correlations bewteen SSO and all these items individually. They were all small, ranging from only `r -1*min_corr` to `r -1*max_corr`. Although we hypothesized that SSO would mediate the relationship between partner's other objectification and interaction authenticity, after finding no relationship bewteen SSO and authencity, we also tested if the partner's other objectification had a direct effect on authenticity, but this effect was not significant, `r report_stats(apim2.b, 3)` (nor was the total effect of partner's other objectification on authenticity, `r report_stats(apim2.c, 2)`).

@@

summary(apim5.c)
```
Lastly, althought there was no evidence that SSO was related to interaction authenticity in the current sample, we tested if interaction authenticity (composite of nine items) had effects on cognitive performace, career aspirations, and relationship agency, as it did in @garcia2016objectification. We again used MLM and thus, these efefcts were tested in three separate models. There was no significant effect of interaction authenticity on cognitive performance, `r report_stats(apim3, 2)`, but authenticity was significantly positively related to both career aspirations, `r report_stats(apim5, 2)`, and relationships agency, `r report_stats(apim4, 2)`. There was no direct effect of SSO on cognitive performance, `r report_stats(apim3.b, 3)`, and no direct effect of partner's other objectificaiton on cognifitve performance, `r report_stats(apim3.c, 4)`. There was no direct effect of SSO on career aspirations, `r report_stats(apim5.b, 3)`, and no direct effect of partner's other objectificaiton on career aspirations, `r report_stats(apim5.c, 4)`. There was no direct effect of SSO on relationship agency, `r report_stats(apim4.b, 3)`, and no direct effect of partner's other objectificaiton on relationship agency, `r report_stats(apim4.c, 4)`.

@@

##Sexism and Sexual Discrimination
Sexism has recently been found to be related to the objectification of women, but not men, for both men and women [@harsey2020men]. Empirical evidence illustrates how women continue to be objects of interpersonal discrimination and expereince daily sexist hassles [@swimetal]. One form of interpersonal discrimination women face is the process by which their whole being is viewed as a collection of sexualized body parts valued predominantly for commodification, a phenomena termed sexual objectification [@Bartky]. Sexual objectification occurs with both “endless variety and monotonous similarity,” and is thus mediated by unique combinations of race, ethnicity, sexuality, age, and class [@fredricksonetal2011; @rubin1975, cited in @frasernicholson1989, p. 28]. Amid such heterogeneity though, “having a reproductively mature female body” proposed by @robertsfredrickson is likely to create a shared vulnerability to sexual objectification and a variety of shared negative experiences as a result.

@@

###Measurement
Also, future experiments or longitudinal studies should explore the external validity of the notions of self-objectification and how the operalization of self-objectification may be improved.

>>>>

@robertsfredrickson Objectification Theory, suggests that women are objectified in interpersonal encounters, and that the effect might be strongest from men, but possibly also from women. Psychological researchers studying the sexual objectification of women, and the resulting self-objectification that follows, have in recent years explored the interpersonal process of the objectification - self-objectification link. For example, in a face-to-face mixed-gender dyadic study, @garcia2016objectification found that men's reported objectification of their female interaction partner during an interaction was associated with increased self-objectification reported by their partner. @gervais2020social has reviewed the literature on interpersonal self-objectification and has organized the the process in a theoretical model called the SIMO, but this model is only relevant for mixed-gender encounters. Although there is ample evidence that women objectify other women [@loughnan2015exploring; @puvia2013being], there is currently no studies investigating the process of interpersonal objectification same-gender interactions. Does objectification by a female interaction partner have the same downstream negative consequences as being objectified by a male interaction partner? For example, perhaps women could be amplifying each others' self-objectification as they get ready to go out for the night, *before* the theoretical encounters with men outlined in @gervais2020social occur.

@@

SIMO model and literature review [@gervais2020social]. Studies have shown that within social encounters women are gazed at more than men [@briton1995, CITE saguy Skype study, calogero video walking study], often times feel ‘looked at’ within interpersonal interactions [@argyle1969], and will more than likely internalize the objectifying gaze on physical self [@young1979]. Moreover, perhaps the most adverse effect of objectifying treatment is that it effectively socializes girls and women to treat themselves as objects to be looked at and evaluated, an effect termed self-objectification  [@Bartky; @berger1972; @fredricksonetal1998]. We don't know about women-on-women interpersonal objectification. What might be different?

Self-objectification has been found to occur after a mere relationship prime among women [@sanchez2008romance] because in our culture women need to look attractive to obtain and maintain successful relationship, thus, the "college relationships condition may heighten self-objectification and the evaluation of other women in a sexualized way. Researchers have also studied interpersonal objectification and self-objectification in romantic relationships [@strelan2018birds]. Evidence is mixed, and women can sometimes have positive outcomes from appearance valuation within the context of an established romantic relationship and they experience less SSO after appearance comments within romantic relationships [@meltzer2020women], less is known about first dates. But this literature is also only about heterosexual relationships. A recent study has investigated the effect of self-objectification on the reduction of relationship building skills in general (including same-sex friendships) [CITE PWQ paper I reviewed, Yoder editor]. Empirical evidence reveals that objectification manifests through inauthenticity in romantic relationships [@brunelletal2010], adverse attitudes in regard to career aspirations, and a decrease in concentration and impairment in female cognitive performance [@quinnetal]. Thus, there is evidence that the expereince of state self-objectification in mixed-gender contexts (stranger and romantic) and within the self in any moment (trying on a swimsuit) has negative consequences, but what about in the context of interaction with other women?

There is some evidence that reduced authenticity is a consequence of self-objectification in the moment [@garcia2016objectification; again CITE PWQ paper I reviewed, Yoder editor]. This link has been justified by the lierature on stigmatized-stigmatizer interactions, viewing the experience of being objectified in an interaction as an identity threat situation. When a woman is objectified by a man, and then experiences self-objectification, this intergroup encounter might trigger identitiy threat, but is the same negative consequence present when a woman is objectified by another woman? woman-woman interpersonal objectification processes might diverge from this process. One the one hand, she may not be having the same negative consequences the cascade from situations that threated group-based identity [@deaux1987putting; @dovidio2006nonverbal; @hebl2005promoting], but feeling like a body, rather than a full human, in an interaction may be enough to reduce feelings of authenticity and social competnce, regardless of the gender of the objectifyer [@tolman2006looking; PWQ paper I reviewed, Yoder editor].

@@

The decision was made to ask even heterosexual women to judge their women partners as potential romantic partners. We felt that this prompt would keep the study closest to a replication of the previous @garcia2016objectification version of the study. Past research has found that women, largely due to an increase self-objectifying state, are indeed able to evaluate other women's potential as romantic partners--indeed, women may be unfortunately used to thinking this way about themselves, and we suspect other women.

@@

#Smith--external and internal
smith_funcTSO_alpha <- psych::alpha(select(smith_final, tso_1, tso_2, tso_4, tso_7, tso_9), check.keys=TRUE)$total[2]
smith_physTSO_alpha <- psych::alpha(select(smith_final, tso_3, tso_5, tso_6, tso_8, tso_10), check.keys=TRUE)$total[2]
```
Trait self-objectification (TSO) was assessed using the Self-Objectification Questionnaire [@fredricksonetal1998; @nollfredrickson1998], which evaluates the extent to which individuals view their bodies in observable versus non-observable ways. The questionnaire asked participants to rank order both appearance and functional aspects of their bodies, from 1 (least important) to 10 (most important), with respect to physical self-concepts. Of the ten body attributes, five of the items were appearance-based (weight, sex appeal, physical attractiveness, firm/sculpted muscles and body measurements), and five of the items were competence-based (strength, physical coordination, energy level, health and physical fitness). Difference scores were computed by subtracting the sum of the 5 functional aspects/competence attributes (e.g., health, strength) from the sum of the 5 physical self-concepts/appearance attributes (e.g., physical attractiveness, weight), and all measures were multiplied by -1, as was done in @garcia2016objectification, so that positive scores indicated greater TSO.

@@

```{r 2016figure, include=TRUE, echo=FALSE, results='asis', out.width="400px", fig.cap="Path Analysis Model from Garcia et al. (2016) study with distinguishable dyads."}
include_graphics("2016_figure.jpg")
```
This study sought to replicate the results of @garcia2016objectification’s study, done with male-female pairs, which used a dyadic path analysis to detect whether partners’ objectification of one another affected state self-objectification (SSO). See Figure\ \@ref(fig:2016figure) for the results of the analysis from this previous study. We hope to investigate how the central effects found in the previous study relate to interactions between two women. Specifically, we are interested in testing the relationship between state-other objectification and SSO, and how SSO in turn, affects feelings of inauthenticity during the interaction. In addition, we will also test if the effect of other-objectification in an interaction on SSO is only present for those women who are high in trait self-objectification, as in @garcia2016objectification. Further, we will investigate the relationships between experiencing interaction inauthenticity and relationship agency, career aspirations, and cognitive performance.

While @garcia2016objectification used dyadic path analysis, we will conduct our dyadic analyses using multilevel modeling. Dyadic analyses for distinguishable dyads (e.g., mixed-gender interacting pairs) is more natural in Structural Equation Modeling (SEM) than it is for indistinguishable dayds (e.g., same-gender interacting pairs) [@ledermann2017analyzing; @garcia2015moderation]. One reason for this asymmetry is that, due to the arbitrary distinctions made between "partner 1" and "partner 2" in indistinguishable dyads, many estimates need to be fixed to be equal (i.e., paths, variances, covariances, endogenous intercepts, and exogenous means) for indistinguishable dyads but these equality constraints should not then be considered in the degrees of freedom calculations for fit estimations [@olsen2006structural]. Further, @olsen2006structural detail how a new independence model and the corresponding fit measure should be re-calculated for indistinguishable dyads models. The current study uses dyadic multilevel modeling (MLM) to test all relationships and mediation patterns. The online supplementary materials contains analysis using SEM. See @ledermann2017analyzing for a more complete discussion of the considerations for using SEM versus MLM for dyadic analysis.

Testing the @garcia2016objectification model on the current, same-gender, sample, involves using the Actor-Partner Independence Model (APIM) approach for each outcome variable (i.e., endogenous variable in Figure\ \@ref(fig:2016figure)). Thus, we ran five APIM's to test all the hypothesized relationships. See Figure\ \@ref(fig:apim) for a basic APIM model. The APIM includes effects due to one's own, as well as one's partner's, predictor variables ($X$'s) on the one's own outcome variable ($Y$). Unlike the original @garcia2016objectification study, our study deals with indistinguishable dyads, meaning the designation of who is designated as "actor" and who is designated as "partner" is arbitrary. Recall that the indistinguishable nature of the dyads in the current study led us to choose the MLM approach over SEM. These analyses are considered exploratory, given the lack of prior research theorizing about these linkages.

@@

round(corr.test(select(filter(clarksmith_women, smith_or_clark == "Clark"), 
                 a2pSOO_A, a2pSOO_P, TSO_A, 
                 howbody_avg_A,
                 authenticity_A, 
                 trigram_score_A, careerasp_A, relag_A))$r, 3)
```
The correlations between study variables is similar across samples. The reliabilities for the study scales were also equivalent. (Note that the two samples were two small to conduct formal measurement equivalence tests for scales.) There are no statistically significant differences between samples in demographics including age, STATS, and ethnicity, STATS.

@@

summary(apim1)
summary(apim1.b)
```
The most important finding from @garcia2016objectification was the significant partner effect of other objectification and SSO (specifically men's objectification of women and women's SSO). As expected, the partner effect of other-objectification on SSO in the current all-women sample was statistically significant, `r report_stats(apim1, 3)`, replicating @garcia2016objectification's finding. One's own other objectification had no effect on SSO, `r report_stats(apim1, 2)`. Contrary to past finding however, there was no statistically significant interaction of partner's other objectification and the person's trait self-objectification on SSO, `r report_stats(apim1, 5)`. There was also no significant main effect of trait self-objectification on SSO `r report_stats(apim1.b, 4)`.

@@

summary(apim2.c)
```
Contrary to expectations, there was no significant effect of SSO on interaction authenticity, although the estimate of this effect was in the hypothesized negative direction, `r report_stats(apim2, 2)`. Because authenticity was a composite score of 9 items, two of which were interaction specific authenticity items, we also estimated the pairwise correlations between SSO and all these items individually. They were all small, ranging from only `r -1*min_corr` to `r -1*max_corr`. Although we hypothesized that SSO would mediate the relationship between partner's other objectification and interaction authenticity, after finding no relationship between SSO and authenticity, we also tested if the partner's other objectification had a direct effect on authenticity, but this effect was not significant, `r report_stats(apim2.b, 3)` (nor was the total effect of partner's other objectification on authenticity, `r report_stats(apim2.c, 2)`).

@@

summary(apim5.c)
```
Lastly, although there was no evidence that SSO was related to interaction authenticity in the current sample, we tested if interaction authenticity (composite of nine items) had effects on cognitive performance, career aspirations, and relationship agency, as it did in @garcia2016objectification. We again used MLM and thus, these effects were tested in three separate models. There was no significant effect of interaction authenticity on cognitive performance, `r report_stats(apim3, 2)`, but authenticity was significantly positively related to both career aspirations, `r report_stats(apim5, 2)`, and relationships agency, `r report_stats(apim4, 2)`. There was no direct effect of SSO on cognitive performance, `r report_stats(apim3.b, 3)`, and no direct effect of partner's other objectification on cognitive performance, `r report_stats(apim3.c, 4)`. There was no direct effect of SSO on career aspirations, `r report_stats(apim5.b, 3)`, and no direct effect of partner's other objectification on career aspirations, `r report_stats(apim5.c, 4)`. There was no direct effect of SSO on relationship agency, `r report_stats(apim4.b, 3)`, and no direct effect of partner's other objectification on relationship agency, `r report_stats(apim4.c, 4)`.

@@

##Sexism and Sexual Discrimination
Sexism has recently been found to be related to the objectification of women, but not men, for both men and women [@harsey2020men]. Empirical evidence illustrates how women continue to be objects of interpersonal discrimination and experience daily sexist hassles [@swimetal]. One form of interpersonal discrimination women face is the process by which their whole being is viewed as a collection of sexualized body parts valued predominantly for commodification, a phenomena termed sexual objectification [@Bartky]. Sexual objectification occurs with both “endless variety and monotonous similarity,” and is thus mediated by unique combinations of race, ethnicity, sexuality, age, and class [@fredricksonetal2011; @rubin1975, cited in @frasernicholson1989, p. 28]. Amid such heterogeneity though, “having a reproductively mature female body” proposed by @robertsfredrickson is likely to create a shared vulnerability to sexual objectification and a variety of shared negative experiences as a result.

@@

###Measurement
Also, future experiments or longitudinal studies should explore the external validity of the notions of self-objectification and how the operationalization of self-objectification may be improved.
